SafeNet: A European database of plant-pollinator interactions

### 2024-02-20 Manuscript/Ms.qmd
3rd paragraph... some progress

@@

Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018]. While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators. However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016]. Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide. Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024]. A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., sampling effort or methodology), which hampers our understanding of plant-pollinator interactions. Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.

3rd paragraph

Low connectance?

@@

Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level. The primary focus of the pollinator taxonomic groups of this database are Hymenoptera, Diptera, Lepidoptera and Coleoptera.

@@

This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent. This dataset consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**). The dataset contains a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species. The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**). These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).

>>>>

#Main orders percentage
o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
main_orders = data %>% 
filter(Pollinator_order %in% o)
main_orders_percent = nrow(main_orders) / nrow(data) * 100

@@

Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018]. While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators. However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016]. Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide. Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024]. A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., habitat type or floral availability), which hampers our understanding of plant-pollinator interactions. Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.

The interactions between different plant and pollinator species within a community form complex networks. Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of the interaction ecology of the different taxonomic groups across space [REF] and environmental gradients [@saunders2023]. In addition, environmental gradients has served as useful tool to explore the impacts of climate on phenological change [REF:Byers 2017]. Importantly, macro-ecological analyses of plant-pollinator networks rarely account for differences in sampling effort [REF] or sampling methodology [REF], ignoring important components that can largely influence our view of plant-pollinator networks [REF]. In addition, despite how relevant are interactions to understand plant-pollinator communities, most communities remain under-sampled [REFERENCES] due to XXXXXXXSpecies occurrences?. Thus, large scale datasets of plant-pollinator interactions so researchers can account for differences across studies...

@@

Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level. The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe. These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, which comprise almost the totality of recorded interactions in the SafeNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).

SafeNet provides a comprehensive overview of plant-pollinator interactions at European level

@@

This European metaweb consist of phyto-centric species level networks of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent. This dataset contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**). The dataset includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species. The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**). These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
